Surface-driven registration method for the structure-informed segmentation of diffusion MR images\tnoteref{t1}

### 2015-10-26 Publications/2015-NeuroImage/04-methods.tex
v1.3.0rc

  \begin{equation}
  P(U \mid M, \omegaset ) = \frac{P(M \mid U,\omegaset )\, P(U)}{P(M)},
  \label{eq:regseg-bayes_rule}
  \end{equation}
  where $P(M \mid U,\omegaset)$ is the data likelihood.
Because $\omegaset$ are mapped by $U$, we simplify
  $P(U, \omegaset) = P(U) \implies P(M \mid U,\omegaset) = P(M \mid U)$.
The best estimate $\hat{U}$ then satisfies the maximum a posteriori criterion
  \citep{bishop_pattern_2006} and it aligns $\gammaset_R$ into $M$.
First, we assume independence between pixels, and thus we break down the
  global data likelihood into a product of pixel-wise conditional probabilities:

  \begin{equation}
  P(M \mid U) = \underset{l}{\prod} \underset{\vec{r}\in \Omega_l}{\prod}
    P\left( \vec{f}' \mid U \right),
  \label{eq:regseg-bayes_aposteriori}
  \end{equation}
  where $\vec{f}' = M(\vec{r}')$ is the feature vector at the displaced
  position $\vec{r}'$ \eqref{eq:regseg-transform} in the moving image.
For convenience and because it has been shown to be an appropriate approximation
  \citep{leemput_automated_1999,cuadra_comparison_2005}, we introduce two assumptions for each
  region $\Omega_l$:
  1) the features are i.i.d.; and 2) they can be modeled by multivariate normal
  distributions with the parameters $\lbrace \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} \rbrace$
  for each region $\Omega_l$.
\newcomment[RV\#1(C.4)]{%
Even though the features being segmented are not generally i.i.d., the spatial interdependency of
  voxels is implicitly supported by the piecewise smooth partition of the space \omegaset{}.
In \autoref{fig:regseg-model} it is shown that marginal distributions of data can be approximated by
  multivariate normal distributions.} Therefore,

  \begin{align}
  P( M \mid U) &= \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod}
  \mathcal{N} ( \vec{f} \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} ) =   \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod} \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{l}\right|}}\,{e^{\left(-\frac{1}{2}
  \mdist{f'}{l} \right)}},
  \label{eq:regseg-pdf}
  \end{align}
  using $\mdist{f}{l}$ to denote the squared \emph{Mahalanobis distance} of $\vec{f}$ with respect
  to the descriptors of region $l$ as
  $\mdist{f}{l} = (\vec{f} - \boldsymbol{\mu}_l)^T \, {\boldsymbol{\Sigma}_l}^{-1} \, (\vec{f} - \boldsymbol{\mu}_l)$.
$C$ is the number of channels comprised in the image $M$.
In fact, the projection of $\gammaset_R$ onto $M$ is an implicit segmentation model, for which
  the covariance matrix $\boldsymbol{\Sigma}_l$ of each region is minimized.
\newcomment[RV\#1(C.2)]{%
In \autoref{fig:regseg-model},
  this minimization is illustrated for the segmentation of the \gls*{fa} and the \gls*{adc} maps
  of one subject.}

\begin{figure}
\includestandalone[width=\linewidth]{figure02-model}
% \includegraphics[width=\linewidth]{figure04A}
% \includegraphics[width=\linewidth]{figure04B}
% \includegraphics[width=\linewidth]{figure04C}
\caption{%
\newcomment{%
Evolution of the segmentation model defined by the homogeneous regions $\Omega_l$
  for a real dataset.
A) shows the initialization of the model, when the anatomically-correct surfaces are not
  mapped into the \gls*{dmri} space.
The panel shows the resulting conditional distributions of segmenting the (distorted) \gls*{fa} and
  \gls*{adc} maps with the original (undistorted) surfaces.
Panel B shows the equivalent information, using the surfaces with perfect matching with the distorted space
  of \gls*{dmri}.
Since the surfaces are now aligned in $M$, the spreads of the conditional distributions for each region are
  reduced with respect to those in A.
Panel C, after \regseg{} registration, shows how the covariance matrices $\boldsymbol{\Sigma}_l$
  are minimized with respect to panel A, and the resulting segmentation is close
  to the theoretical situation presented in panel B.
The left (large) plot in each panel represents a kernel density estimate of the location and spread of
  each region $\Omega_l$ in the bivariate feature space of $M$, which comprises
  the \gls*{fa} and \gls*{adc} maps.
At the top and the right margins, the marginal distributions of $\Omega_l$ are also plotted for
  the \gls*{fa} and \gls*{adc}, respectively.
On the right, the sampled distribution of each region is shown in small scatter plot, in contrast to
  the complete bivariate data.}
}\label{fig:regseg-model}
\todo[inline]{RV\#1(C.5)}
\end{figure}

@@

The adaptation of this protocol to the simulated phantoms and real data is explained in the
  following sections.
\newcomment[RV\#2(C.20)]{%
A first proof of concept to demonstrate \regseg{} in phantoms with simple
  geometries is introduced, using $U_\text{true}$ without directional restrictions.
Then, a more realistic environment with $U_\text{true}$ restricted to the
  \gls*{pe} axis of real \gls*{dmri} datasets is presented.}

@@

\newcomment[RV\#1(C.6)]{\paragraph*{Segmentation model and settings}
The segmentation model of the phantoms is implicitly defined in each shape.
All phantoms comprehend an inner surface enclosing a uniform \gls*{wm}-like region,
  and an outer surface wrapping a \gls*{gm}-like layer.
The outside is filled with uniform background (see \autoref{fig:regseg-evphantoms}).
All the experimental settings used for the phantoms are made available in
  a unique configuration file\footnote{\url{https://github.com/oesteban/RegSeg/blob/master/Scripts/pyacwereg/data/regseg_default.json}}.}

@@

\begin{figure*}
\includegraphics[width=\linewidth]{figure03}
\caption{Experimental workflow employed to process real data from the \acrfull*{hcp}.
  1) $\gammaset_R$ were extracted from the anatomical reference (\gls*{t1} image).
  2) For use as the ground truth, we generated a plausible synthetic distortion $U_\text{true}$
    from the field map with \eqref{eq:regseg-fieldmap}.
  3) The \gls*{dmri} data were warped using $U_\text{true}$ to reproduce the effects of real
    susceptibility-derived distortions.
  Target diffusion scalars (\gls*{fa} and \gls*{adc}) were computed with the distorted data and
    stacked to feed the multivariate input required by \regseg{}.
  4) The method was run to obtain $U_\text{test} = \hat{U}_\text{true}$, i.e., the estimate of
    the ground-truth deformation.
  5) The results were evaluated visually and quantitatively.
  \newcomment{%
  The arrows point to edges in the target images (yellow arrows for \gls*{fa}, blue for \gls*{adc}) that
    should be aligned with a surface, showing how distortion breaks the mapping with the structural space in which
    the contours are defined.}}\label{fig:regseg-evworkflows}
  \todo[inline]{RV\#2(C.23)}%
\end{figure*}

\paragraph*{Data}
To evaluate \regseg{} using real \gls*{dmri} data obtained from human brains,
  we collected 16 subjects from the \gls*{hcp} database.
The original acquisitions are released within ``unprocessed'' packages, whereas
  the ``minimally preprocessed'' packages contain the corresponding images after
  some processing (correction for several artifacts, brain-extraction, spatial
  normalization, etc.).
We refer the reader to \citep{essen_human_2012} for exact details of the acquisition
  parameters and \citep{glasser_minimal_2013} for the preprocessing issues.
These datasets comprise a large set of images, including \gls*{t1}, \gls*{t2}, and
  multi-shell \gls*{dmri} images.
\newcomment[RV\#1(C.7)]{%
Since we obtained the \gls*{dmri} data from the minimally preprocessed package, these
  images are corrected for \gls*{epi} distortions and spatially normalized in
  \gls*{t1} space.}

@@

\paragraph*{\Regseg{} settings} %
\newcomment[RV\#1(C.5)]{%
\Regseg{} accepts an affine mapping from surface-space to the \gls*{dmri} data as initialization.
However, the images provided by the \gls*{hcp} are already spatially normalized.
Therefore, the initial estimation of distortion is zero in this experiment.}
\newcomment[RV\#1(C.20)]{%
Since the distortion $U_\text{true}$ is aligned along the \gls*{pe} direction ($y$-axis in our settings),
  \regseg{} was configured to allow nonzero displacements only on that corresponding direction.}
\newcomment[RV\#1(C.6)]{%
For the experiments on real data, \regseg{} established a multi-resolution pyramid of
  B-Spline basis, with control points distributed on grids of the following spacings:
  40$\times$100$\times$40 [mm] for the first (coarser level), \isores[mm]{30} for the second,
  and 20$\times$30$\times$10 [mm] in the third.
Only the first and second levels included Gaussian smoothing of the target image ($\sigma$ = [2.0, 0.5] mm,
  respectively).
The actual choices of the parameter settings are publicly distributed with the source code for the
  experiments\footnote{\url{https://github.com/oesteban/RegSeg/tree/master/Scripts/pyacwereg/data/regseg_hcp.json}}.}
These settings were obtained manually, driven by the feedback obtained from the post-registration
  convergence reports (like that found in \suppl{section S1.3}).
We released \regseg{} along with the tool to generate such convergence reports.

\paragraph*{Cross-comparison and settings} %
A dual workflow to the general evaluation framework used for \regseg{} (\autoref{fig:regseg-evworkflows}),
  was employed to integrate the alternate \gls*{t2b} registration scheme.
We reproduced the solution and settings provided with \emph{ExploreDTI}
  \citep{leemans_exploredti_2009}, which is a widely used toolkit for tractography analysis of
  \gls*{dti}.
\emph{ExploreDTI} internally employs \emph{elastix} \citep{klein_elastix_2010} to
  perform registration.
\newcomment[RV\#2(C.21)]{%
The deformation field is correspondingly restricted to the \gls*{pe} direction.
Settings used with \emph{elastix} are conveniently available in its format\footnote{\url{https://github.com/oesteban/RegSeg/blob/master/Scripts/pyacwereg/data/t2b_elastix_y.txt}}.}

>>>>

  \begin{equation}
  P(U \mid M, \omegaset ) = \frac{P(M \mid U,\omegaset )\, P(U)}{P(M)},
  \label{eq:regseg-bayes_rule}
  \end{equation}
  where $P(M \mid U,\omegaset)$ is the data likelihood.
Because $\omegaset$ are mapped by $U$, we simplify
  $P(U, \omegaset) = P(U) \implies P(M \mid U,\omegaset) = P(M \mid U)$.
The best estimate $\hat{U}$ then satisfies the maximum a posteriori criterion
  and it aligns $\gammaset_R$ into $M$.
First, we assume independence between pixels, and thus we break down the
  global data likelihood into a product of pixel-wise conditional probabilities:

  \begin{equation}
  P(M \mid U) = \underset{l}{\prod} \underset{\vec{r}\in \Omega_l}{\prod}
    P\left( \vec{f}' \mid U \right),
  \label{eq:regseg-bayes_aposteriori}
  \end{equation}
  where $\vec{f}' = M(\vec{r}')$ is the feature vector at the displaced
  position $\vec{r}'$ \eqref{eq:regseg-transform} in the moving image.
For convenience and because it has been shown to be an appropriate approximation
  \citep{leemput_automated_1999,cuadra_comparison_2005}, we introduce two assumptions for each
  region $\Omega_l$:
  1) the features are i.i.d.; and 2) they can be modeled by multivariate normal
  distributions \citep{esteban_mbis_2014} with parameters
  $\lbrace \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} \rbrace$
  for each region $\Omega_l$.
\revcomment[RV\#1(C.4)]{%
Even though the features being segmented are not generally i.i.d., the spatial interdependency of
  voxels is implicitly supported by the piecewise smooth partition of the space \omegaset{}.
In \autoref{fig:regseg-model} it is shown that marginal distributions of data can be approximated by
  multivariate normal distributions.} Therefore,

  \begin{align}
  P( M \mid U) &= \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod}
  \mathcal{N} ( \vec{f} \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} ) =   \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod} \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{l}\right|}}\,{e^{\left(-\frac{1}{2}
  \mdist{f'}{l} \right)}},
  \label{eq:regseg-pdf}
  \end{align}
  using $\mdist{f}{l}$ to denote the squared \emph{Mahalanobis distance} of $\vec{f}$ with respect
  to the descriptors of region $l$ as
  $\mdist{f}{l} = (\vec{f} - \boldsymbol{\mu}_l)^T \, {\boldsymbol{\Sigma}_l}^{-1} \, (\vec{f} - \boldsymbol{\mu}_l)$.
$C$ is the number of channels comprised in the image $M$.
In fact, the projection of $\gammaset_R$ onto $M$ is an implicit segmentation model, for which
  the covariance matrix $\boldsymbol{\Sigma}_l$ of each region is minimized.
\revcomment[RV\#1(C.2)]{%
In \autoref{fig:regseg-model},
  this minimization is illustrated for the segmentation of the \gls*{fa} and the \gls*{adc} maps
  of one subject.}

\begin{figure*}
\includestandalone[width=\linewidth]{figure02-model}
% \includegraphics[width=\linewidth]{figure04A}
% \includegraphics[width=\linewidth]{figure04B}
% \includegraphics[width=\linewidth]{figure04C}
\caption{%
\revcomment{%
Evolution of the segmentation model defined by the homogeneous regions $\Omega_l$
  for a real dataset.
A) shows the initialization of the model, when the anatomically-correct surfaces are not
  mapped into the \gls*{dmri} space.
The panel shows the resulting conditional distributions of segmenting the (distorted) \gls*{fa} and
  \gls*{adc} maps with the original (undistorted) surfaces.
Panel B shows the equivalent information, using the surfaces with perfect matching with the distorted space
  of \gls*{dmri}.
Since the surfaces are now aligned in $M$, the spreads of the conditional distributions for each region are
  reduced with respect to those in A.
Panel C, after \regseg{} registration, shows how the covariance matrices $\boldsymbol{\Sigma}_l$
  are minimized with respect to panel A, and the resulting segmentation is close
  to the theoretical situation presented in panel B.
The left (large) plot in each panel represents a kernel density estimate of the location and spread of
  each region $\Omega_l$ in the bivariate feature space of $M$, which comprises
  the \gls*{fa} and \gls*{adc} maps.
At the top and the right margins, the marginal distributions of $\Omega_l$ are also plotted for
  the \gls*{fa} and \gls*{adc}, respectively.
On the right, the sampled distribution of each region is shown in small scatter plot, in contrast to
  the complete bivariate data.}
}\label{fig:regseg-model}
\todo[inline]{RV\#1(C.5)}
\end{figure*}

@@

The adaptation of this protocol to the simulated phantoms and real data is explained in the
  following sections.
\revcomment[RV\#2(C.20)]{%
A first proof of concept to demonstrate \regseg{} in phantoms with simple
  geometries is introduced, using $U_\text{true}$ without directional restrictions.
Then, a more realistic environment with $U_\text{true}$ restricted to the
  \gls*{pe} axis of real \gls*{dmri} datasets is presented.}

@@

\revcomment[RV\#1(C.6)]{\paragraph*{Segmentation model and settings}
The segmentation model of the phantoms is implicitly defined in each shape.
All phantoms comprehend an inner surface enclosing a uniform \gls*{wm}-like region,
  and an outer surface wrapping a \gls*{gm}-like layer.
The outside is filled with uniform background (see \autoref{fig:regseg-evphantoms}).
All the experimental settings used for the phantoms are made available in
  a unique configuration file\footnote{\url{https://github.com/oesteban/RegSeg/blob/master/Scripts/pyacwereg/data/regseg_default.json}}.}

@@

\begin{figure*}
\includegraphics[width=\linewidth]{figure03}
\caption{Experimental workflow employed to process real data from the \acrfull*{hcp}.
  1) $\gammaset_R$ were extracted from the anatomical reference (\gls*{t1} image).
  2) For use as the ground truth, we generated a plausible synthetic distortion $U_\text{true}$
    from the field map with \eqref{eq:regseg-fieldmap}.
  3) The \gls*{dmri} data were warped using $U_\text{true}$ to reproduce the effects of real
    susceptibility-derived distortions.
  Target diffusion scalars (\gls*{fa} and \gls*{adc}) were computed with the distorted data and
    stacked to feed the multivariate input required by \regseg{}.
  4) The method was run to obtain $U_\text{test} = \hat{U}_\text{true}$, i.e., the estimate of
    the ground-truth deformation.
  5) The results were evaluated visually and quantitatively.
  \revcomment{%
  The arrows point to edges in the target images (yellow arrows for \gls*{fa}, blue for \gls*{adc}) that
    should be aligned with a surface, showing how distortion breaks the mapping with the structural space in which
    the contours are defined.}}\label{fig:regseg-evworkflows}
  \todo[inline]{RV\#2(C.23)}%
\end{figure*}

\paragraph*{Data}
To evaluate \regseg{} using real \gls*{dmri} data obtained from human brains,
  we collected 16 subjects from the \gls*{hcp} database.
The original acquisitions are released within ``unprocessed'' packages, whereas
  the ``minimally preprocessed'' packages contain the corresponding images after
  some processing (correction for several artifacts, brain-extraction, spatial
  normalization, etc.).
We refer the reader to \citep{essen_human_2012} for exact details of the acquisition
  parameters and \citep{glasser_minimal_2013} for the preprocessing issues.
These datasets comprise a large set of images, including \gls*{t1}, \gls*{t2}, and
  multi-shell \gls*{dmri} images.
\revcomment[RV\#1(C.7)]{%
Since we obtained the \gls*{dmri} data from the minimally preprocessed package, these
  images are corrected for \gls*{epi} distortions and spatially normalized in
  \gls*{t1} space.}

@@

\paragraph*{\Regseg{} settings} %
\revcomment[RV\#1(C.5)]{%
\Regseg{} accepts an affine mapping from surface-space to the \gls*{dmri} data as initialization.
However, the images provided by the \gls*{hcp} are already spatially normalized.
Therefore, the initial estimation of distortion is zero in this experiment.}
\revcomment[RV\#1(C.20)]{%
Since the distortion $U_\text{true}$ is aligned along the \gls*{pe} direction ($y$-axis in our settings),
  \regseg{} was configured to allow nonzero displacements only on that corresponding direction.}
\revcomment[RV\#1(C.6)]{%
For the experiments on real data, \regseg{} established a multi-resolution pyramid of
  B-Spline basis, with control points distributed on grids of the following spacings:
  40$\times$100$\times$40 [mm] for the first (coarser level), \isores[mm]{30} for the second,
  and 20$\times$30$\times$10 [mm] in the third.
Only the first and second levels included Gaussian smoothing of the target image ($\sigma$ = [2.0, 0.5] mm,
  respectively).
The actual choices of the parameter settings are publicly distributed with the source code for the
  experiments\footnote{\url{https://github.com/oesteban/RegSeg/tree/master/Scripts/pyacwereg/data/regseg_hcp.json}}.}
These settings were obtained manually, driven by the feedback obtained from the post-registration
  convergence reports (like that found in \suppl{section S1.3}).
We released \regseg{} along with the tool to generate such convergence reports.

\paragraph*{Cross-comparison and settings} %
A dual workflow to the general evaluation framework used for \regseg{} (\autoref{fig:regseg-evworkflows}),
  was employed to integrate the alternate \gls*{t2b} registration scheme.
We reproduced the solution and settings provided with \emph{ExploreDTI}
  \citep{leemans_exploredti_2009}, which is a widely used toolkit for tractography analysis of
  \gls*{dti}.
\emph{ExploreDTI} internally employs \emph{elastix} \citep{klein_elastix_2010} to
  perform registration.
\revcomment[RV\#2(C.21)]{%
The deformation field is correspondingly restricted to the \gls*{pe} direction.
Settings used with \emph{elastix} are conveniently available in its format\footnote{\url{https://github.com/oesteban/RegSeg/blob/master/Scripts/pyacwereg/data/t2b_elastix_y.txt}}.}

### 2015-10-26 Publications/2015-NeuroImage/06-discussion.tex
v1.3.0rc

An important antecedent of \regseg{} is \emph{bbregister} \citep{greve_accurate_2009}.
The tool has been widely adopted as the standard registration method to be used along with the \gls*{epi}
  correction of choice.
It implements a linear mapping, and uses 3D active contours \emph{with edges} to
  search for intensity boundaries in the \lowb{} image.
The active contours are initialized using surfaces extracted from the \gls*{t1} using
  \emph{FreeSurfer} \citep{fischl_freesurfer_2012}: the \emph{pial} surface (exterior of the
  cortical \gls*{gm}) and the \emph{white} surface (the \gls*{wm}/\gls*{gm} interface).
The \lowb{} image only includes a detectable frontier for the pial surface, and thus
  \emph{bbregister} is limited to aligning the cortical layer in this
  application.
To overcome the problem of nonlinear distortions, \emph{bbregister} excludes the
  regions that are typically warped from the boundary search.
Indeed, the distortion must be addressed separately because it is not supported by
  the affine transformation model.
The deformation model of \regseg{} is nonlinear and the surfaces are active
  contours \emph{without edges} \citep{chan_active_2001}.
The contours are without edges because the \gls*{fa} and \gls*{adc} maps do
  not present steep image gradients (edges) but the anatomy can be identified
  by looking for piece-wise smooth homogeneous regions.

@@

We verified that precise segmentation and registration of a set of surfaces into multivariate
  data is possible on digital phantoms.
We randomly deformed four different phantom models to mimic three homogeneous regions
  (\gls*{wm}, \gls*{gm}, and \acrlong*{csf}) and we used them to simulate \gls*{t1} 
  and \gls*{t2} images at two resolution levels.
After registration with \regseg{}, we measured the Hausdorff distance between the
  projected contours obtained using the ground-truth warping and our estimates.
We concluded that the errors were significantly lower than the voxel resolution.
We also assessed the 95\% \gls*{ci}, which yielded an aggregate interval of
  0.64--0.66 [mm] for the low resolution phantoms (2.0 mm isotropic voxel) and
  0.34--0.38 [mm] for the high resolution phantoms (1.0 mm isotropic).
Therefore, we also concluded that the error was bounded above by half of the
  voxel spacing.
\newcomment[RV\#1(C.8)]{%}
The distributions of errors along surfaces vary importantly depending on the shape of the
  phantom (see \autoref{fig:regseg-phantom}B).
The misregistration error of the ``gyrus'' phantom showed a much lower spread than that
  for the other shapes.
We argue that the symmetry of those other shapes posed difficulties in driving the contours
  towards the appropriate region and producing some ``sliding'' effect between the
  faces of the surfaces and their ground-truth position.
This effect should not be present in real datasets, thanks to the very convoluted cortical
  layer, and the directional restriction of distortion.}

\todo{real data}
This application is necessary in the extraction of structural connectivity from
  \gls*{dmri}, and has been typically solved in a two-step approach.
First, the \glspl*{dwi} are corrected for \emph{\gls*{epi} distortions} by estimating
  the nonlinear-deformation field from extra MR acquisitions
  \citep{jezzard_correction_1995,chiou_simple_2000,cordes_geometric_2000,
  kybic_unwarping_2000}.
Second, mapping the structural information from the corresponding \gls*{t1} image
  using a linear registration tool like \emph{bbregister} \citep{greve_accurate_2009}.
numerous attempts to delineate the structures of interest in diffusion
  space have been proposed.
Current activity on improving correction methods \citep{irfanoglu_drbuddi_2015} and
  segmenting the \gls*{dmri} in its native space \citep{jeurissen_tissuetype_2015}
  proofs the open interest of the application.
\todo{real data old}
We designed \regseg{} as a method for segmenting \gls*{dmri} data by exploiting the
  anatomical information extracted from a structural image (typically a \gls*{t1} image)
  of the subject.
Current pipelines using whole-brain tractography usually solve this problem with a two-step approach.
First, the images are corrected for nonlinear distortions using auxiliary acquisitions
  such as fieldmaps \citep{jezzard_correction_1995}, \glspl*{dwi} with reversed \gls*{pe}
  blips \citep{chiou_simple_2000}, or \gls*{t2} images \citep{kybic_unwarping_2000}.
Second, the segmentation is projected from a reference \gls*{t1} image using linear
  registration \citep{greve_accurate_2009}.
\Regseg{} addresses this joint problem in a single step and it does not require any additional
  acquisition other than the minimal protocol using only \gls*{t1} and \gls*{dmri} images.
This situation is found commonly in historical datasets.
Moreover, since the structural information is projected into the native space of \gls*{dmri}
  reconstruction and tractography can be performed without resampling data to an undistorted
  space.

We evaluated \regseg{} in a real environment using the experimental framework presented
  in \autoref{fig:regseg-evworkflows}.
We processed 16 subjects from the \gls*{hcp} database using both \regseg{}
  and an in-house replication of the \acrfull*{t2b} method.
\Regseg{} obtained very high accuracy, with an aggregate 95\% \gls*{ci} of 0.56--0.66 [mm], which was
  below the pixel size of 1.25 mm.
The misregistration error that remained after \regseg{} was significantly lower ($p <$ 0.01) than the
  error corresponding to the \gls*{t2b} correction according to Kruskal-Wallis H-tests
  (\autoref{tab:results_real}).
Visual inspections of all the results (\suppl{section S5}) and the violin plots in
  \autoref{fig:regseg-results_real} confirmed that \regseg{} performed better than the \gls*{t2b} method
  in our settings.
We carefully configured the \gls*{t2b} method using the same algorithm and the
  same settings employed in a widely-used tool.
However, cross-comparison experiments are prone to so-called \emph{instrumentation bias}
  \citep{tustison_instrumentation_2013}.
Therefore, these results do not prove that \regseg{} \emph{is better than} \gls*{t2b}.
Our results suggest that \regseg{} is a reliable option in this application field.
In addition, the \gls*{t2b} may introduce an additional (and small) error during the necessary
  registration of \gls*{t2} in the \gls*{t1} space.

\newcomment[RV\#1(C.3)]{%
\paragraph*{Contributions}\label{sec:regseg-related_work} %

}

\newcomment[RV\#1(C.5)]{%
Finally, we also contribute proposing a piecewise-smooth segmentation model defined by
  a selection of nested surfaces to partition the multispectral space
  comprehending the \gls*{fa} and the \gls*{adc} maps and ultimately identify these
  structures in \gls*{dmri} space.}
\newcomment[New]{%
With the definition of appropriate models, \regseg{} may be applied in other fields like
  neonatal brain image segmentation in longitudinal \gls*{mri} studies like \citep{shi_neonatal_2010}
  in which the surfaces obtained in a mature timepoint of the brain can be retrospectively
  propagated to the initial timepoints, regardless the changes of the contrast and spatial
  development between timepoints.}

>>>>

An important antecedent of \regseg{} is \emph{bbregister} \citep{greve_accurate_2009}.
The tool has been widely adopted as the standard registration method to be used along with the \gls*{epi}
  correction of choice.
It implements a linear mapping, and uses 3D active contours \emph{with edges} to
  search for intensity boundaries in the \lowb{} image.
The active contours are initialized using surfaces extracted from the \gls*{t1} using
  \emph{FreeSurfer} \citep{fischl_freesurfer_2012}: the \emph{pial} surface (exterior of the
  cortical \gls*{gm}) and the \emph{white} surface (the \gls*{wm}/\gls*{gm} interface).
The \lowb{} image only includes a detectable frontier for the pial surface, and thus
  \emph{bbregister} is limited to aligning the cortical layer in this
  application.
To overcome the problem of nonlinear distortions, \emph{bbregister} excludes the
  regions that are typically warped from the boundary search.
Indeed, the distortion must be addressed separately because it is not supported by
  the affine transformation model.
Conversely, the deformation model of \regseg{} is nonlinear and the surfaces are active
  contours \emph{without edges} \citep{chan_active_2001}.
The contours are without edges because the \gls*{fa} and \gls*{adc} maps do
  not present steep image gradients (edges) but the anatomy can be identified
  by looking for piece-wise smooth homogeneous regions.

@@

We verified that precise segmentation and registration of a set of surfaces into multivariate
  data is possible on digital phantoms.
We randomly deformed four different phantom models to mimic three homogeneous regions
  (\gls*{wm}, \gls*{gm}, and \acrlong*{csf}) and we used them to simulate \gls*{t1} 
  and \gls*{t2} images at two resolution levels.
After registration with \regseg{}, we measured the Hausdorff distance between the
  projected contours obtained using the ground-truth warping and our estimates.
We concluded that the errors were significantly lower than the voxel size.
We also assessed the 95\% \gls*{ci}, which yielded an aggregate interval of
  0.64--0.66 [mm] for the low resolution phantoms (2.0 mm isotropic voxel) and
  0.34--0.38 [mm] for the high resolution phantoms (1.0 mm isotropic).
Therefore, we also concluded that the error was bounded above by half of the
  voxel size.
\revcomment[RV\#1(C.8)]{%}
The distributions of errors along surfaces vary importantly depending on the shape of the
  phantom (see \autoref{fig:regseg-phantom}B).
The misregistration error of the ``gyrus'' phantom showed a much lower spread than that
  for the other shapes.
We argue that the symmetry of those other shapes posed difficulties in driving the contours
  towards the appropriate region and producing some ``sliding'' effect between the
  faces of the surfaces and their ground-truth position.
This effect should not be present in real datasets, thanks to the very convoluted cortical
  layer, and the directional restriction of distortion.}

\todo{real data}
Once the method was validated, we applied it in the task of identifying or incorporating
  structural information in 16 real \gls*{dmri} datasets.
The segmentation of some anatomical structures in the diffusion dataset is necessary in
  current methods involved in the connectome extraction, such as the structure-informed 
  reconstruction of \gls*{dmri} data \citep{jeurissen_multitissue_2014,daducci_accelerated_2015},
  the anatomically constrained tractography \citep{smith_anatomicallyconstrained_2012},
  and the imposition of the cortical parcellation mapped from the \gls*{t1} image 
  \citep{hagmann_mapping_2008}.
The problem was firstly addressed using image segmentation approaches in the native diffusion
  space, without definitive and compelling results.
With the introduction of retrospective correction methods for the \emph{\gls*{epi} distortions}
  and image registration approaches, the task has been typically solved in a two-step approach.
First, the \glspl*{dwi} are corrected for \emph{\gls*{epi} distortions} by estimating
  the nonlinear-deformation field from extra MR acquisitions
  \citep{jezzard_correction_1995,chiou_simple_2000,cordes_geometric_2000,
  kybic_unwarping_2000}.
Second, mapping the structural information from the corresponding \gls*{t1} image
  using a linear registration tool like \emph{bbregister} \citep{greve_accurate_2009}.
The current activity on improving correction methods \citep{irfanoglu_drbuddi_2015} and
  the comeback of segmentation of \gls*{dmri} in its native space 
  \citep{jeurissen_tissuetype_2015} proof the open interest of this application.

\Regseg{} addresses this joint problem in a single step and it does not require any additional
  acquisition other than the minimal protocol using only \gls*{t1} and \gls*{dmri} images.
This situation is found commonly in historical datasets.
Moreover, since the structural information is projected into the native space of \gls*{dmri},
  reconstruction and tractography can be performed without resampling data to an undistorted
  space.
We evaluated \regseg{} in a real environment using the experimental framework presented
  in \autoref{fig:regseg-evworkflows}.
We processed 16 subjects from the \gls*{hcp} database using both \regseg{}
  and an in-house replication of the \acrfull*{t2b} method.
\Regseg{} obtained very high accuracy, with an aggregate 95\% \gls*{ci} of 0.56--0.66 [mm], which was
  below the pixel size of 1.25 mm.
The misregistration error that remained after \regseg{} was significantly lower ($p <$ 0.01) than the
  error corresponding to the \gls*{t2b} correction according to Kruskal-Wallis H-tests
  (\autoref{tab:results_real}).
Visual inspections of all the results (\suppl{section S5}) and the violin plots in
  \autoref{fig:regseg-results_real} confirmed that \regseg{} achieved higher accuracy
  than the \gls*{t2b} method in our settings.
We carefully configured the \gls*{t2b} method using the same algorithm and the
  same settings employed in a widely-used tool for \gls*{dmri} processing.
However, cross-comparison experiments are prone to the so-called \emph{instrumentation bias}
  \citep{tustison_instrumentation_2013}.
Therefore, these results did not prove that \regseg{} \emph{is better than} \gls*{t2b},
  but indicated that \regseg{} is a reliable option in this application field.
\revcomment[RV\#1(C.5)]{%
Finally, we also contributed proposing a piecewise-smooth segmentation model defined by
  a selection of nested surfaces to partition the multispectral space
  comprehending the \gls*{fa} and the \gls*{adc} maps and ultimately identify these
  structures in \gls*{dmri} space.
We also demonstrated the smoothness of the objective function on five of the real datasets
  (\suppl{Figure S2}), taking advantage of the directional restriction of possible distortions.
However, convergence of \regseg{} can be severely affected when surfaces are dense enough.
Using the digital phantoms, we decimated the surfaces with large factors and
  effectively displaced the minimum of the shape-gradients from the minimum of the
  objective function.
Thus, the surfaces must be densely sampled to enable the use of the presented method.
}%

Beyond the presented application on \gls*{dmri} data, \regseg{} can be indicated in situations
  where there are precise surfaces delineating the structure, a target multivariate
  image in which the surfaces must be fitted, and the mapping between the surfaces and
  the volume encodes a relevant physiological information, such as the normal/abnormal
  development or the macroscopic dynamics of organs and tissues.
For instance, \regseg{} may be applied in fields like neonatal brain image segmentation
  in longitudinal MRI studies of the early developmental patterns as \citep{shi_neonatal_2010}.
In these studies, the surfaces obtained in a mature timepoint of the brain are retrospectively
  propagated to the initial timepoints, regardless the changes of the contrast and spatial
  development between them.
More generally, \regseg{} may also be applied on the personalized study of longitudinal alteration
  of the brain using multispectral images, for instance in the case of traumatic brain
  injury \citep{irimia_structural_2014} or in the development of brain tumors
  \citep{weizman_semiautomatic_2014}.
